Syntax Is All You Need: A Universal-Language Approach to Mutant Generation

### 2024-02-15 paper/main.tex
maybe camera ready

%%
%% This is file `sample-acmsmall-conf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `acmsmall-conf')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-acmsmall-conf.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%%
%% The first command in your LaTeX source must be the \documentclass
%% command.
%%
%% For submission and review of your manuscript please change the
%% command to \documentclass[manuscript, screen, review]{acmart}.
%%
%% When submitting camera ready or to TAPS, please change the command
%% to \documentclass[sigconf]{acmart} or whichever template is required
%% for your publication.
%%
%%
\documentclass[acmsmall,screen,review,anonymous]{acmart}

@@

@@

\section{A Universal Language Syntax-Based Approach}

The key insight of this paper is that \emph{most mutation operators
  proposed in the literature do not in fact require parsing of source
  code.}  Consider, for example, one of the most commonly used
mutation operators, replacement of arithmetic operations.  We do not
need to parse a program containing the string {\tt ``x + y''} in order to
replace that string with {\tt ``x - y''}.  Instead, this change is
guaranteed to be effected if we simply search the program, represented
as a string, for all occurrences of  {\tt ``+''}, and produce for each
such occurrence, a mutant where tha character is replaced by {\tt
  ``-''}.  Many languages express arithmetic as infix operators between
expressions; if expressed syntactically, mutations for them can be expressed
once, trivially implementing a large fraction of the classic Mothra~\cite{offutt1996experimental}
mutation operators.

The obvious objection to this simplistic approach is that unless we
parse the code to identify arithmetic expressions, some of these
replacements will be invalid.  E.g., in a C++ program we will produce
mutants like {\tt x-+} replacing {\tt x++}.  However, most of these
instances can be avoided by slightly more judicious choice of search
target, e.g., instances of {\tt ``+''} where the preceeding or following
character is not another {\tt ``+''}.  This may still in some cases produce
invalid mutants, but rejecting such mutants is easy, since they will fail
to compile.  Of course, we pay the price of
compiling each mutant at generation
time.

@@

Our tool has been available as an open source project on GitHub (with
over 100 stars to date) and
available for installation via Python pip for
some time, has received contributions from users not involved in
our project, and has been adopted by at least one flagship project
(over 19K stars) as
its officially recommended mutation tool.  We omit details for
purposes of blinding.

@@

    \begin{tabular}{l|l|l|r}
    %\toprule
    \toprule
    \textbf{Language} & \textbf{Tool} & \textbf{URL} & \textbf{LOC}  \\
    \midrule
    \multirow{2}{*}{Python}  & Mutmut & https://github.com/boxed/mutmut & 3870  \\
        & cosmic-ray & https://github.com/sixty-north/cosmic-ray & 4599 \\ 
    \midrule
    \multirow{2}{*}{Java}  & PIT & https://github.com/hcoles/pitest & 59577  \\
        & LittleDarwin & https://github.com/aliparsai/LittleDarwin & 22359 \\ 
    \midrule
        Rust & Cargo-Mutants & https://github.com/sourcefrog/cargo-mutants & 7020 \\
    \midrule
        C++ & Dextool & https://github.com/joakim-brannstrom/dextool & 38611 \\
    \midrule
        ALL & Our Tool & removed for blinding & 2244 \\
    \bottomrule
    \end{tabular}
\end{table}

@@

Hariri et al. compared C mutation approaches at the source and
compiler IR levels~\cite{CompareSrcBinary} and found that overall
source level mutation was better, producing fewer mutants overall, but
matching the IR approach in the important measures of surface and
minimal mutants and overall mutation score.  Numerous studies compare
Java mutation tools~\cite{MajorPIT,gopinath2017does}, including a
recent article for a more general audience in Communications of the
ACM~\cite{CommACMJavaTool} (perhaps showing the growing interest in
practical mutation testing), which showed that users
considered active maintenance, support for a variety of testing
frameworks, and support for recent Java versions as the most important
features in a Java mutation tool.  The approach proposed in this paper
by its nature tends to promise all three of these key factors without
imposing burdens on maintainers. More recently, there has been some effort to create a 
language agnostic mutation testing tool~\cite{wodeltest}.  However, this tool uses
Wodel, a new (and hard-to-read, compared to Comby) DSL for expressing operators, and requires users to
define a set of operators for each language, without sharing  
of rules.  An
  additional paper on this general topic is a short tool paper for an initial
  prototype of our current tool, omitted for blinding.

%Finally, the practical use of mutation testing at Google argues that
%support for a variety of languages is critical.  As noted above, we
%cover the set of languages and operators used in Google's internal
%mutation testing tools~\cite{PetrovicMutationGoogle}, but in an open
%source, easily extensible tool offering a large number of additional
%languages and operators.

>>>>

%%
%% This is file `sample-acmsmall-conf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `acmsmall-conf')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-acmsmall-conf.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%%
%% The first command in your LaTeX source must be the \documentclass
%% command.
%%
%% For submission and review of your manuscript please change the
%% command to \documentclass[manuscript, screen, review]{acmart}.
%%
%% When submitting camera ready or to TAPS, please change the command
%% to \documentclass[sigconf]{acmart} or whichever template is required
%% for your publication.
%%
%%
\documentclass[acmsmall]{acmart}

@@

\author{Sourav Deb}
\affiliation{\institution{Northern Arizona University}
\country{United States}}

\author{Kush Jain}
\affiliation{\institution{Carnegie Mellon University}
\country{United States}}

\author{Rijnard van Tonder}
\affiliation{\institution{Mysten Labs}
\country{United States}}

\author{Claire Le Goues}
\affiliation{\institution{Carnegie Mellon University}
\country{United States}}

\author{Alex Groce}
\affiliation{\institution{Northern Arizona University}
\country{United States}}

@@

       Our tool implementing these ideas, the \emph{universalmutator}, is available on GitHub at
       \url{https://github.com/agroce/universalmutator}, and can be
       installed via Python pip.    We
       welcome contributions and improvement suggestions.  Our tool is
       the official recommended tool for mutation of Solidity programs (\url{https://docs.soliditylang.org/en/latest/resources.html}).

\section{A Universal-Language Syntax-Based Approach}

The key insight of this paper is that \emph{most mutation operators
  proposed in the literature do not in fact require parsing of source
  code.}  Consider, for example, one of the most commonly used
mutation operators, replacement of arithmetic operations.  We do not
need to parse a program containing the string {\tt ``x + y''} in order to
replace that string with {\tt ``x - y''}.  Instead, this change is
guaranteed to be effected if we simply search the program, represented
as a string, for all occurrences of  {\tt ``+''}, and produce for each
such occurrence, a mutant where tha character is replaced by {\tt
  ``-''}.  Many languages express arithmetic as infix operators between
expressions; if expressed syntactically, mutations for them can be expressed
once, trivially implementing a large fraction of the classic Mothra~\cite{offutt1996experimental}
mutation operators.  In a \emph{syntactical} (but not semantic) sense, we claim, there is almost a ``universal'' (very
abstract, never implemented) programming language, of which almost all
particular programming languages can be treated as mere dialects.

The obvious objection to this simplistic approach is that unless we
parse the code to identify arithmetic expressions, some of these
replacements will be invalid; the universal language is not actually real!  E.g., in a C++ program we will produce
mutants like {\tt x-+} replacing {\tt x++}.  However, most of these
instances can be avoided by slightly more judicious choice of search
target, e.g., instances of {\tt ``+''} where the preceeding or following
character is not another {\tt ``+''}.  This may still in some cases produce
invalid mutants, but rejecting such mutants is easy, since they will fail
to compile.  Of course, we pay the price of
compiling each mutant at generation
time.

@@

%Our tool has been available as an open source project on GitHub (with
%over 100 stars to date) and
%available for installation via Python pip for
%some time, has received contributions from users not involved in
%our project, and has been adopted by at least one flagship project
%(over 19K stars) as
%its officially recommended mutation tool.  We omit details for
%purposes of blinding.

@@

    \begin{tabular}{l|l|l|r}
    %\toprule
    \toprule
    \textbf{Language} & \textbf{Tool} & \textbf{URL} & \textbf{LOC}  \\
    \midrule
    \multirow{2}{*}{Python}  & Mutmut & https://github.com/boxed/mutmut & 3870  \\
        & cosmic-ray & https://github.com/sixty-north/cosmic-ray & 4599 \\ 
    \midrule
    \multirow{2}{*}{Java}  & PIT & https://github.com/hcoles/pitest & 59577  \\
        & LittleDarwin & https://github.com/aliparsai/LittleDarwin & 22359 \\ 
    \midrule
        Rust & Cargo-Mutants & https://github.com/sourcefrog/cargo-mutants & 7020 \\
    \midrule
        C++ & Dextool & https://github.com/joakim-brannstrom/dextool & 38611 \\
    \midrule
        ALL & Universalmutator & https://github.com/agroce/universalmutator & 2244 \\
    \bottomrule
    \end{tabular}
\end{table}

@@

Hariri et al. compared C mutation approaches at the source and
compiler IR levels~\cite{CompareSrcBinary} and found that overall
source level mutation was better, producing fewer mutants overall, but
matching the IR approach in the important measures of surface and
minimal mutants and overall mutation score.  Numerous studies compare
Java mutation tools~\cite{MajorPIT,gopinath2017does}, including a
recent article for a more general audience in Communications of the
ACM~\cite{CommACMJavaTool} (perhaps indicating a growing interest in
practical mutation testing), which showed that users
considered active maintenance, support for a variety of testing
frameworks, and support for recent Java versions as the most important
features in a Java mutation tool.  The approach proposed in this paper
by its nature tends to promise all three of these key factors without
imposing burdens on maintainers. More recently, there has been some effort to create a 
language agnostic mutation testing tool~\cite{wodeltest}.  However, this tool uses
Wodel, a new (and hard-to-read, compared to Comby) DSL for expressing operators, and requires users to
define a set of operators for each language, without sharing  
of rules.  Groce et al. originally presented an initial
  prototype of our approach in a short tool paper, lacking parser-parser combinator
  support or a non-trivial empirical evaluation~\cite{regexpMut}.  
  The universalmutator
  was used in previous work proposing mutation-based evaluation and
  improvement of static analysis tools~\cite{StatMut}, as a basis
  for the mutation approach (using Comby) adopted in a paper on
  improving turn-key compiler fuzzing~\cite{NoFuss}, and as a basis
  for incorporating mutation generation in fuzzing~\cite{FirstFuzz},
  as well as in other work simply using it as a mutation tool.
%Finally, the practical use of mutation testing at Google argues that
%support for a variety of languages is critical.  As noted above, we
%cover the set of languages and operators used in Google's internal
%mutation testing tools~\cite{PetrovicMutationGoogle}, but in an open
%source, easily extensible tool offering a large number of additional
%languages and operators.
